---
title: "Advanced Usage of latexSymb: Building Complex Mathematical Documents"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage of latexSymb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(latexSymb)
data(common)
attach(common)
```

## Introduction

This vignette demonstrates advanced techniques for using `latexSymb` to create complex mathematical documents. We'll explore how to organize your code effectively, create reusable functions, and build sophisticated mathematical expressions.

## Code Organization: The Two-File Pattern

When working with complex mathematical documents, it's beneficial to separate your symbol definitions and helper functions from your main document content. This pattern offers several advantages:

1. **Cleaner Documents**: Your .Rmd file remains focused on content rather than setup
2. **Reusability**: Symbol definitions can be shared across multiple documents
3. **Maintainability**: Changes to symbols or functions are centralized
4. **Readability**: Mathematical logic is easier to follow without setup code

### Example: Setting Up a Symbol Library

Create a separate .R file for your symbols, for instance:

```{r}
# my_symbols.R
library(latexSymb)
data(common)
attach(common)

# Define document-specific symbols
pt <- lsymb("x")
pt_prime <- lsymb("x'")
metric_space <- lsymb("\\mathcal{X}")
measure <- lsymb("\\mu")

# Create specialized functions
dist <- function(x, y) lsymb("d") * pths(x * comma * y)
```

Then, use a separate .Rmd to write the body of your document. In that file, set:

```
knitr::opts_chunk$set(results = "asis", echo = FALSE)
source("my_symbols.R")
```

to make sure that the LaTeX code gets correctly rendered.

## Creating Powerful Helper Functions

You can use `latexSymb` to handle complicated mathematical expressions. The following patterns can be helpful.

### Pattern 1: Building Complex Expressions

Create functions that compose to build sophisticated expressions:

```{r}
# Expectation operator
exp_val <- function(x) lsymb("\\mathbb{E}") * sqbr(x)

sq <- function(x) pths(x)^2

# Absolute value
abs <- function(x) lsymb("\\abs{", x, "}")

# Indicator function
indic <- function(condition) {
  lsymb("\\mathbbm{1}") |> under(br(condition))
}

# Composed example
X <- lsymb("X")
Y <- lsymb("Y")
```

Then you can create expressions like `exp_val(sq(abs(X - Y))) |> il()` to obtain `r exp_val(sq(abs(X - Y))) |> il()` and `exp_val(X * indic(Y > 0)) |> il()` for `r xp_val(X * indic(Y > 0)) |> il()`.

### Pattern 2: Multi-line Equation Builders

This pattern uses lists and the `lenv()` function. Each element of the list corresponds to a row of your displayed expression:

```{r}
# Build a proof step by step
proof_steps <- list(
  ruler * dist(x, y) * leq * dist(x, z) + dist(z, y) * endl,
  ruler * thus * dist(x, y) - dist(x, z) * leq * dist(z, y)
) |>
  lenv("align*", rows = _)
```

`r proof_steps`


## Complex Example: Statistical Proof

Here's how these techniques combine in practice:

```{r}
# Setup
sample_size <- lsymb("n")
observation <- lsymb("X") |> under(i)
sample_mean <- lsymb("\\bar{X}")
conv_distr <- lsymb("\\overset{d}{\\rightarrow}")
sqrt <- function(x) lsymb("\\sqrt{", x, "}")

# Build a Central Limit Theorem statement
clt_statement <- list(
  ruler * sqrt(sample_size) * (pths(sample_mean - mu) / si) *
    conv_distr * lsymb("N(0,1)") * endl,
  ruler * lsymb("\\text{where }") *
    sample_mean * eq * (1 / sample_size) * Sum(observation, from = i * eq * 1, to = sample_size)
) |>
  lenv("align*", rows = _)
```

`r clt_statement`

## Conclusion

By organizing your code into separate files, creating reusable functions, and following consistent patterns, you can efficiently create complex mathematical documents with `latexSymb`. The key is to:

1. **Separate concerns**: Keep symbol definitions in .R files
2. **Build abstractions**: Create functions for repeated patterns
3. **Compose functions**: Build complex expressions from simple parts
